---
title: "Total_goals_in_likely_draw_matches"
author: "Kutálek Jiří"
date: "2024-02-09"
output:
  html_document:
    theme: default
    smart: no
    toc_depth: 0
    highlight: tango
    df_print: default
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, tidy = FALSE, warning = TRUE, results = 'markup', comment = NA, background = '#FFCC00') # adjust according to your preferences
```

```{r}
library(tidyr)
library(dplyr)
library(agricolae)
library(ggplot2)
library(base)
```

### Read data
```{r}
src_basename = "matches_02_04"
src_filename = "matches_02_04.csv"

dat_raw <- read.csv2(file = paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\", src_filename), header = TRUE, sep = ",")
```

### Adjust data
- convert to `factors`
- convert `floats` (**Expected Goals** and **Odds**) to `numerics`
- convert other numerical values to `integers`
- convert date & time to `datetime`
- replace conflicting `:` symbol with `_`
- add attribute for **total goals** in match
```{r}
dat <- dat_raw |> mutate(
  across(
    c(team_home, team_away, country, competition, season, referee, neutral_field, finished, no_spectators), factor))
dat <- dat |> mutate(
  across(
    where(is.character) & (starts_with("odd") | starts_with("expected")), as.numeric))
dat <- dat |> mutate(
  across(
    where(is.character) & !c(id, date_time), as.integer))

dat$date_time <- as.POSIXct(dat$date_time, format="%Y-%m-%d %H:%M", tz="UTC")
dat <- dat |> mutate(
  competition = gsub(":", "_", competition)
)

dat <- dat |> mutate(
  total_goals = goals_home + goals_away
)
```

### Add column for **match group** `(country, competition, season)`
```{r}
dat <- dat |> mutate(
    match_group = factor(paste0(country, "_", competition, "_", season)))
```

### Add columns for **margins**
- four columns in total - **start**/**end** ("long time before match start"/"right before amtch start") and **Tipsport**/**Fortuna**
- `NA_real_` if zero or invalid odd values
```{r}
dat <- dat |> mutate(
  margin_tipsport_start = ifelse(
    is.na(odd_tipsport_1_start) | is.na(odd_tipsport_0_start) | is.na(odd_tipsport_2_start) | 
      odd_tipsport_1_start <= 0   | odd_tipsport_0_start <= 0   | odd_tipsport_2_start <= 0,
    NA_real_,
    (1/odd_tipsport_1_start + 1/odd_tipsport_0_start + 1/odd_tipsport_2_start) - 1),
  margin_tipsport_end = ifelse(
    is.na(odd_tipsport_1_end) | is.na(odd_tipsport_0_end) | is.na(odd_tipsport_2_end) | 
      odd_tipsport_1_end <= 0   | odd_tipsport_0_end <= 0   | odd_tipsport_2_end <= 0,
    NA_real_,
    (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end) - 1),
  margin_fortuna_start = ifelse(
    is.na(odd_fortuna_1_start) | is.na(odd_fortuna_0_start) | is.na(odd_fortuna_2_start) | 
      odd_fortuna_1_start <= 0   | odd_fortuna_0_start <= 0   | odd_fortuna_2_start <= 0,
    NA_real_,
    (1/odd_fortuna_1_start + 1/odd_fortuna_0_start + 1/odd_fortuna_2_start) - 1),
  margin_fortuna_end = ifelse(
    is.na(odd_fortuna_1_end) | is.na(odd_fortuna_0_end) | is.na(odd_fortuna_2_end) | 
      odd_fortuna_1_end <= 0   | odd_fortuna_0_end <= 0   | odd_fortuna_2_end <= 0,
    NA_real_,
    (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end) - 1)
)

# Print out average margin (start and end)
print(paste0("Average margin by Tipsport for win/draw/loose is: ",
             mean(dat$margin_tipsport_start, na.rm = TRUE), ", ",
             mean(dat$margin_tipsport_end, na.rm = TRUE)))
print(paste0("Average margin by Fortuna for win/draw/loose is: ",
             mean(dat$margin_fortuna_start, na.rm = TRUE), ", ",
             mean(dat$margin_fortuna_end, na.rm = TRUE)))

last_N_columns = 4
random_rows <- sample(nrow(dat), 10) # randomly take 10 rows
dat[random_rows, c(1, (ncol(dat) - (last_N_columns-1)) : ncol(dat))]
```

### Add columns for win/draw/lose **probabilities without margins** (only **end**, not **start**)
- `prob_without_margin = prob_with_margin - prob_with_margin * (margin / sum_of_odds)`
- (`sum_of_odds = 1 + margin`)
- **two possibilities** of how to distirbute the total margin (**A** is the same as **C**, but is **C** is **simpler** and thus, **better !!!**)
```{r}
# A. Distribute the total margin PROPORTIONALLY to the 1/0/2 odds

# dat <- dat |> mutate(
#   prb_nomarg_1_tips_end = (1/odd_tipsport_1_end) - (1/odd_tipsport_1_end) * (margin_tipsport_end / # (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end)),
#   prb_nomarg_0_tips_end = (1/odd_tipsport_0_end) - (1/odd_tipsport_0_end) * (margin_tipsport_end / # (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end)),
#   prb_nomarg_2_tips_end = (1/odd_tipsport_2_end) - (1/odd_tipsport_2_end) * (margin_tipsport_end / # (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end)),
#   prb_nomarg_1_fort_end = (1/odd_fortuna_1_end) - (1/odd_fortuna_1_end) * (margin_fortuna_end / # (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end)),
#   prb_nomarg_0_fort_end = (1/odd_fortuna_0_end) - (1/odd_fortuna_0_end) * (margin_fortuna_end / # (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end)),
#   prb_nomarg_2_fort_end = (1/odd_fortuna_2_end) - (1/odd_fortuna_2_end) * (margin_fortuna_end / # (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end))
# )

# B. Distribute the total margin EQUALLY between the 1/0/2 odds (one third to each)

# dat <- dat |> mutate(
#   prb_nomarg_1_tips_end = (1/odd_tipsport_1_end) - (margin_tipsport_end / 3),
#   prb_nomarg_0_tips_end = (1/odd_tipsport_0_end) - (margin_tipsport_end / 3),
#   prb_nomarg_2_tips_end = (1/odd_tipsport_2_end) - (margin_tipsport_end / 3),
#   prb_nomarg_1_fort_end = (1/odd_fortuna_1_end) - (margin_fortuna_end / 3),
#   prb_nomarg_0_fort_end = (1/odd_fortuna_0_end) - (margin_fortuna_end / 3),
#   prb_nomarg_2_fort_end = (1/odd_fortuna_2_end) - (margin_fortuna_end / 3)
# )

# C. Distribute the total margin in the easiest way - divide each 1/odd by total prob. with margin
dat <- dat |> mutate(
  prb_nomarg_1_tips_end = (1/odd_tipsport_1_end) / (margin_tipsport_end + 1),
  prb_nomarg_0_tips_end = (1/odd_tipsport_0_end) / (margin_tipsport_end + 1),
  prb_nomarg_2_tips_end = (1/odd_tipsport_2_end) / (margin_tipsport_end + 1),
  prb_nomarg_1_fort_end = (1/odd_fortuna_1_end) / (margin_fortuna_end + 1),
  prb_nomarg_0_fort_end = (1/odd_fortuna_0_end) / (margin_fortuna_end + 1),
  prb_nomarg_2_fort_end = (1/odd_fortuna_2_end) / (margin_fortuna_end + 1)
)

# Print out average probabilities without margins (end, not start)
print(paste0("Average probabilities without margins by Tipsport for win/draw/loose are: ",
             mean(dat$prb_nomarg_1_tips_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_0_tips_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_2_tips_end, na.rm = TRUE)))
print(paste0("Average probabilities without margins by Fortuna for win/draw/loose are: ",
             mean(dat$prb_nomarg_1_fort_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_0_fort_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_2_fort_end, na.rm = TRUE)))

last_N_columns = 6
random_rows <- sample(nrow(dat), 10) # randomly take 10 rows
dat[random_rows, c(1, (ncol(dat) - (last_N_columns-1)) : ncol(dat))]
```

### Filter and check the calculated probabilities
- filter out matches with `NA` values for probabilities
- check that the probabilities for each match summing to `1`
```{r}
dat_filtered <- dat |> filter(!is.na(prb_nomarg_1_tips_end) & !is.na(prb_nomarg_0_tips_end) & !is.na(prb_nomarg_2_tips_end) & !is.na(prb_nomarg_1_fort_end) & !is.na(prb_nomarg_0_fort_end) & !is.na(prb_nomarg_2_fort_end))
print(paste0("New length of data frame is ", nrow(dat_filtered)))

# Check the probabilities
prob_columns <- dat_filtered[c("prb_nomarg_1_tips_end", "prb_nomarg_0_tips_end", "prb_nomarg_2_tips_end", "prb_nomarg_1_fort_end", "prb_nomarg_0_fort_end", "prb_nomarg_2_fort_end")]

tolerance <- 1e-5
rows_outside_range <- apply(prob_columns, 1, function(x) {
  prob_sum <- sum(x, na.rm = TRUE)
  abs(prob_sum - 1) > tolerance
})

print(paste0("Number of wrongly calculated matches probabilities equals to ", sum(rows_outside_range, na.rm = TRUE)))
rows_indices_outside_range <- which(rows_outside_range)
```

### Add column for **absolute difference** in **winning probability** (without margin) for both teams
```{r}
dat_filtered <- dat_filtered |> mutate(
  diff_prb_nomarg_tips_end = abs(prb_nomarg_1_tips_end - prb_nomarg_2_tips_end),
  diff_prb_nomarg_fort_end = abs(prb_nomarg_1_fort_end - prb_nomarg_2_fort_end)
)

last_N_columns = 2
random_rows <- sample(nrow(dat_filtered), 10) # randomly take 10 rows
dat_filtered[random_rows, c(1, (ncol(dat_filtered) - (last_N_columns-1)) : ncol(dat_filtered))]
```

### **Filter out** matches during the COVID-19 period and those without spectators
```{r}
no_spectators_matches <- dat_filtered |> filter(no_spectators == "True")
print(paste0("New length of data frame is ", nrow(dat_filtered)))

dat_filtered <- dat_filtered |> filter(no_spectators == "False" & (date_time < "2020-03-01 00:00" | date_time > "2021-09-01 00:00"))
print(paste0("New length of data frame is ", nrow(dat_filtered)))
```

### **Selection** of possible draw matches
```{r}
draw_matches_tipsport <- dat_filtered |> filter(prb_nomarg_0_tips_end >= 0.28 &
                                     prb_nomarg_0_tips_end <= 0.35 &
                                     diff_prb_nomarg_tips_end <= 0.15)
draw_matches_fortuna <- dat_filtered |> filter(prb_nomarg_0_fort_end >= 0.28 &
                                     prb_nomarg_0_fort_end <= 0.35 &
                                     diff_prb_nomarg_fort_end <= 0.15)

print(paste0("New length of Tipsport data frame is ", nrow(draw_matches_tipsport)))
print(paste0("New length of Fortuna data frame is ", nrow(draw_matches_fortuna)))

last_N_columns = 8
random_rows <- sample(nrow(draw_matches_tipsport), 10) # randomly take 10 rows
draw_matches_tipsport[random_rows, c(1, (ncol(draw_matches_tipsport) - (last_N_columns-1)) : ncol(draw_matches_tipsport))]

last_N_columns = 8
random_rows <- sample(nrow(draw_matches_fortuna), 10) # randomly take 10 rows
draw_matches_fortuna[random_rows, c(1, (ncol(draw_matches_fortuna) - (last_N_columns-1)) : ncol(draw_matches_fortuna))]
```

### Distribute all the selected matches into bins
- as follows: `[2.7-2.8, 2.8-2.9, ..., 3.4-3.5]`
```{r}
draw_matches_tipsport_cut <- draw_matches_tipsport |> mutate(
  bin = cut(prb_nomarg_0_tips_end, 
            breaks = c(0.28, 0.29, 0.30, 0.31, 0.32, 0.33, 0.34, 0.35),
            labels = c('28-29%', '29-30%', '30-31%', '31-32%', '32-33%', '33-34%', '34-35%'),
            include.lowest = TRUE, right = FALSE)
)
table(draw_matches_tipsport_cut$bin)

draw_matches_fortuna_cut <- draw_matches_fortuna |> mutate(
  bin = cut(prb_nomarg_0_fort_end, 
            breaks = c(0.28, 0.29, 0.30, 0.31, 0.32, 0.33, 0.34, 0.35),
            labels = c('28-29%', '29-30%', '30-31%', '31-32%', '32-33%', '33-34%', '34-35%'),
            include.lowest = TRUE, right = FALSE)
)
table(draw_matches_fortuna_cut$bin)
```

##### Check number of common rows in both resulting data frames (**Tipsport** and **Fortuna**)
- first, calculate portion of similar matches chosen as likely-draw for both Tipsport and Fortuna
- then, calculate number and portion of matches similarly placed in the individual bins for both Tipsport and Fortuna
```{r}
common_rows <- merge(draw_matches_tipsport_cut, draw_matches_fortuna_cut, by = "id")
print(paste0(nrow(common_rows)/nrow(draw_matches_tipsport) * 100, "% of rows in new Tipsport data frame is in the Fortuna data frame as well."))
print(paste0(nrow(common_rows)/nrow(draw_matches_fortuna) * 100, "% of rows in new Fortuna data frame is in the Tipsort data frame as well."))

# Check similarities in the individual bins (more important)
draw_matches_2829_in_both <- common_rows |> filter(bin.x == "28-29%" & bin.y == "28-29%")
draw_matches_2930_in_both <- common_rows |> filter(bin.x == "29-30%" & bin.y == "29-30%")
draw_matches_3031_in_both <- common_rows |> filter(bin.x == "30-31%" & bin.y == "30-31%")
draw_matches_3132_in_both <- common_rows |> filter(bin.x == "31-32%" & bin.y == "31-32%")
draw_matches_3233_in_both <- common_rows |> filter(bin.x == "32-33%" & bin.y == "32-33%")
draw_matches_3334_in_both <- common_rows |> filter(bin.x == "33-34%" & bin.y == "33-34%")
draw_matches_3435_in_both <- common_rows |> filter(bin.x == "34-35%" & bin.y == "34-35%")

total_number_of_rows_similar_for_both_in_all_bins = length(draw_matches_2829_in_both) + 
  length(draw_matches_2930_in_both) + length(draw_matches_3031_in_both) +
  length(draw_matches_3132_in_both) + length(draw_matches_3233_in_both) +
  length(draw_matches_3334_in_both) + length(draw_matches_3435_in_both)

print(paste0("Total number of matches similarly placed in the individual bins for both Tipsport and Fortuna in all bins together is ", total_number_of_rows_similar_for_both_in_all_bins, ", which is ", 
      total_number_of_rows_similar_for_both_in_all_bins / nrow(draw_matches_tipsport) * 100, "% of total number of matches selected for Tipsport draw matches and ", 
      total_number_of_rows_similar_for_both_in_all_bins / nrow(draw_matches_fortuna) * 100, "% of total number of matches selected for Fortuna draw matches."))
```

```{r}
# common_rows <- merge(draw_matches_tipsport_cut, draw_matches_fortuna_cut, by = "id") |>
#   rename(bin = bin.x)
# 
# common_rows |>
#   group_by(bin) |>
#   slice_sample(n = 10, replace = FALSE) |>
#   select(prb_nomarg_0_tips_end.x, prb_nomarg_0_fort_end.x) |>
#   group_map(~ print(.x))
```

## All competitions, all seasons (non-COVID)

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- draw_matches_tipsport_cut |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- draw_matches_tipsport_cut |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (all non-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- draw_matches_fortuna_cut |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- draw_matches_fortuna_cut |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (all non-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

### Bar plots of average goals in each bin of matches (by draw odds without margins)
- **Tipsport**
```{r}
# Sort (asc.)
sorted <- draw_matches_tipsport |> arrange(prb_nomarg_0_tips_end)

# Split into 7 bins
data_bin_7 <- sorted |>
  mutate(draw_odds_bin = cut(prb_nomarg_0_tips_end, breaks = 7)) |>
  group_by(draw_odds_bin) |>
  summarize(average_goals = mean(total_goals, na.rm = TRUE),
            match_count = n(), .groups = 'drop')

# Split into 14 bins
data_bin_14 <- sorted |>
  mutate(draw_odds_bin = cut(prb_nomarg_0_tips_end, breaks = 14)) |>
  group_by(draw_odds_bin) |>
  summarize(average_goals = mean(total_goals, na.rm = TRUE),
            match_count = n(), .groups = 'drop')

# Plot
chart <- ggplot() +
  geom_col(data = data_bin_7, aes(x = draw_odds_bin, y = average_goals), fill = "blue", alpha = 1.0) +
  geom_col(data = data_bin_14, aes(x = draw_odds_bin, y = average_goals), fill = "azure4", alpha = 1.0) +
  geom_text(data = data_bin_7, aes(x = draw_odds_bin, y = average_goals, label = round(average_goals, 2)), 
            position = position_dodge(width = 0.9), vjust = -0.5, color = "blue", fontface = "bold") + # Odds labels
  geom_text(data = data_bin_14, aes(x = draw_odds_bin, y = average_goals, label = round(average_goals, 2)), 
            position = position_dodge(width = 0.9), vjust = -0.5, color = "azure4") + # Odds labels
  geom_text(data = data_bin_7, aes(x = draw_odds_bin, y = 0, label = match_count), 
            vjust = 1.5, color = "blue") + # Match count labels
  geom_text(data = data_bin_14, aes(x = draw_odds_bin, y = 0, label = match_count), 
            vjust = 1.5, color = "azure4") + # Match count labels
  labs(x = "Binned draw odds", y = "Mean total goals", title = "Mean number of goals in matches binned by draw odds without margin (all non-COVID matches with spectators; Fortuna)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, face = "bold"),
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

chart

# Save
filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\bars_by_margins_tipsport.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
# Sort (asc.)
sorted <- draw_matches_fortuna |> arrange(prb_nomarg_0_fort_end)

# Split into 7 bins
data_bin_7 <- sorted |>
  mutate(draw_odds_bin = cut(prb_nomarg_0_fort_end, breaks = 7)) |>
  group_by(draw_odds_bin) |>
  summarize(average_goals = mean(total_goals, na.rm = TRUE),
            match_count = n(), .groups = 'drop')

# Split into 14 bins
data_bin_14 <- sorted |>
  mutate(draw_odds_bin = cut(prb_nomarg_0_fort_end, breaks = 14)) |>
  group_by(draw_odds_bin) |>
  summarize(average_goals = mean(total_goals, na.rm = TRUE),
            match_count = n(), .groups = 'drop')

# Plot
chart <- ggplot() +
  geom_col(data = data_bin_7, aes(x = draw_odds_bin, y = average_goals), fill = "blue", alpha = 1.0) +
  geom_col(data = data_bin_14, aes(x = draw_odds_bin, y = average_goals), fill = "azure4", alpha = 1.0) +
  geom_text(data = data_bin_7, aes(x = draw_odds_bin, y = average_goals, label = round(average_goals, 2)), 
            position = position_dodge(width = 0.9), vjust = -0.5, color = "blue", fontface = "bold") + # Odds labels
  geom_text(data = data_bin_14, aes(x = draw_odds_bin, y = average_goals, label = round(average_goals, 2)), 
            position = position_dodge(width = 0.9), vjust = -0.5, color = "azure4") + # Odds labels
  geom_text(data = data_bin_7, aes(x = draw_odds_bin, y = 0, label = match_count), 
            vjust = 1.5, color = "blue") + # Match count labels
  geom_text(data = data_bin_14, aes(x = draw_odds_bin, y = 0, label = match_count), 
            vjust = 1.5, color = "azure4") + # Match count labels
  labs(x = "Binned draw odds", y = "Mean total goals", title = "Mean number of goals in matches binned by draw odds without margin (all non-COVID matches with spectators; Fortuna)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, face = "bold"),
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

chart

# Save
filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\bars_by_margins_fortuna.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## All competitions, only pre-COVID seasons

### Filter out post-COVID matches too
```{r}
pre_covid_matches_tipsport <- draw_matches_tipsport_cut |> filter(date_time < "2020-03-01 00:00")
print(paste0("New length of data frame is ", nrow(pre_covid_matches_tipsport)))

pre_covid_matches_fortuna <- draw_matches_fortuna_cut |> filter(date_time < "2020-03-01 00:00")
print(paste0("New length of data frame is ", nrow(pre_covid_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- pre_covid_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- pre_covid_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (pre-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_preCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- pre_covid_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- pre_covid_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (pre-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_preCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## All competitions, only post-COVID seasons

### Filter out pre-COVID matches too
```{r}
post_covid_matches_tipsport <- draw_matches_tipsport_cut |> filter(date_time > "2021-09-01 00:00")
print(paste0("New length of data frame is ", nrow(post_covid_matches_tipsport)))

post_covid_matches_fortuna <- draw_matches_fortuna_cut |> filter(date_time > "2021-09-01 00:00")
print(paste0("New length of data frame is ", nrow(post_covid_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- post_covid_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- post_covid_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (post-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_postCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- post_covid_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- post_covid_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (post-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_postCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## Only GREAT competitions, all non-COVID seasons

### Filter out low-level competitions too
- WARNING: More competitions are called "Bundesliga" (Germany and Austria?) or "Super League" (Greece and Switzerland?)
```{r}
top_countries <- c("England", "Germany", "France", "Czech Republic", "Spain", "Italy", "Netherlands", "Portugal", "Belgium", "Denmark", "Turkey", "Slovakia", "Israel", "Croatia", "Greece", "Austria", "Switzerland", "Poland", "Sweden", "Norway")
top_countries

top_competitions <- c("FORTUNA_LIGA", "FNL", "Bundesliga", "2. Bundesliga", "Premier League", "Championship", "Ligue 1", "LaLiga", "Serie A", "Eredivisie", "Liga Portugal", "Jupiler Pro League", "Superliga", "Super Lig", "Serie B", "LaLiga2", "Ligue 2", "League One", "3. Liga", "Nike liga", "Ligat ha'Al", "HNL", "Super League", "Ekstraklasa", "Eliteserien", "Allsvenskan")
top_competitions

great_competitions_matches_tipsport <- draw_matches_tipsport_cut |>
  filter(country %in% top_countries & competition %in% top_competitions)
print(paste0("New length of data frame is ", nrow(great_competitions_matches_tipsport)))

great_competitions_matches_fortuna <- draw_matches_fortuna_cut |>
  filter(country %in% top_countries & competition %in% top_competitions)
print(paste0("New length of data frame is ", nrow(great_competitions_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- great_competitions_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- great_competitions_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="slateblue", alpha=0.85, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (GREAT competitions matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_great_competitions.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- great_competitions_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- great_competitions_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="slateblue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (GREAT competitions matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_great_competitions.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## Only GREAT competitions, only pre-COVID seasons

### Filter out low-level competitions and post-COVID matches too
- WARNING: More competitions are called "Bundesliga" (Germany and Austria?) or "Super League" (Greece and Switzerland?)
```{r}
top_countries <- c("England", "Germany", "France", "Czech Republic", "Spain", "Italy", "Netherlands", "Portugal", "Belgium", "Denmark", "Turkey", "Slovakia", "Israel", "Croatia", "Greece", "Austria", "Switzerland", "Poland", "Sweden", "Norway")
top_countries

top_competitions <- c("FORTUNA_LIGA", "FNL", "Bundesliga", "2. Bundesliga", "Premier League", "Championship", "Ligue 1", "LaLiga", "Serie A", "Eredivisie", "Liga Portugal", "Jupiler Pro League", "Superliga", "Super Lig", "Serie B", "LaLiga2", "Ligue 2", "League One", "3. Liga", "Nike liga", "Ligat ha'Al", "HNL", "Super League", "Ekstraklasa", "Eliteserien", "Allsvenskan")
top_competitions

great_competitions_pre_covid_matches_tipsport <- draw_matches_tipsport_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time < "2020-03-01 00:00")
print(paste0("New length of data frame is ", nrow(great_competitions_pre_covid_matches_tipsport)))

great_competitions_pre_covid_matches_fortuna <- draw_matches_fortuna_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time < "2020-03-01 00:00")
print(paste0("New length of data frame is ", nrow(great_competitions_pre_covid_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- great_competitions_pre_covid_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- great_competitions_pre_covid_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="slateblue", alpha=0.85, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (GREAT competitions pre-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_great_competitions_preCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- great_competitions_pre_covid_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- great_competitions_pre_covid_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="slateblue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (GREAT competitions pre-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_great_competitions_preCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## Only GREAT competitions, only pre-COVID seasons

### Filter out low-level competitions and post-COVID matches too
- WARNING: More competitions are called "Bundesliga" (Germany and Austria?) or "Super League" (Greece and Switzerland?)
```{r}
top_countries <- c("England", "Germany", "France", "Czech Republic", "Spain", "Italy", "Netherlands", "Portugal", "Belgium", "Denmark", "Turkey", "Slovakia", "Israel", "Croatia", "Greece", "Austria", "Switzerland", "Poland", "Sweden", "Norway")
top_countries

top_competitions <- c("FORTUNA_LIGA", "FNL", "Bundesliga", "2. Bundesliga", "Premier League", "Championship", "Ligue 1", "LaLiga", "Serie A", "Eredivisie", "Liga Portugal", "Jupiler Pro League", "Superliga", "Super Lig", "Serie B", "LaLiga2", "Ligue 2", "League One", "3. Liga", "Nike liga", "Ligat ha'Al", "HNL", "Super League", "Ekstraklasa", "Eliteserien", "Allsvenskan")
top_competitions

great_competitions_post_covid_matches_tipsport <- draw_matches_tipsport_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time > "2021-09-01 00:00")
print(paste0("New length of data frame is ", nrow(great_competitions_post_covid_matches_tipsport)))

great_competitions_post_covid_matches_fortuna <- draw_matches_fortuna_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time > "2021-09-01 00:00")
print(paste0("New length of data frame is ", nrow(great_competitions_post_covid_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- great_competitions_post_covid_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- great_competitions_post_covid_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="slateblue", alpha=0.85, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (GREAT competitions post-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_great_competitions_postCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- great_competitions_pre_covid_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- great_competitions_pre_covid_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="slateblue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (GREAT competitions post-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_great_competitions_postCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## Only TOP competitions, all non-COVID seasons

### Filter out low-level competitions too
```{r}
top_countries <- c("England", "Germany", "France", "Czech Republic", "Spain", "Italy", "Netherlands", "Portugal", "Belgium", "Denmark", "Turkey", "Greece", "Switzerland", "Croatia")
top_countries

top_competitions <- c("FORTUNA_LIGA", "Bundesliga", "2. Bundesliga", "Premier League", "Championship", "Ligue 1", "LaLiga", "Serie A", "Eredivisie", "Liga Portugal", "Jupiler Pro League", "Superliga", "Super Lig", "Serie B", "LaLiga2", "Super League", "HNL")
top_competitions

top_competitions_matches_tipsport <- draw_matches_tipsport_cut |>
  filter(country %in% top_countries & competition %in% top_competitions)
print(paste0("New length of data frame is ", nrow(top_competitions_matches_tipsport)))

top_competitions_matches_fortuna <- draw_matches_fortuna_cut |>
  filter(country %in% top_countries & competition %in% top_competitions)
print(paste0("New length of data frame is ", nrow(top_competitions_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- top_competitions_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- top_competitions_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="darkorchid", alpha=0.85, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (TOP competitions matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_top_competitions.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- top_competitions_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- top_competitions_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="darkorchid", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (TOP competitions matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_top_competitions.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## Only TOP competitions, only pre-COVID seasons

### Filter out low-level competitions and post-COVID matches too
```{r}
top_countries <- c("England", "Germany", "France", "Czech Republic", "Spain", "Italy", "Netherlands", "Portugal", "Belgium", "Denmark", "Turkey", "Greece", "Switzerland", "Croatia")
top_countries

top_competitions <- c("FORTUNA_LIGA", "Bundesliga", "2. Bundesliga", "Premier League", "Championship", "Ligue 1", "LaLiga", "Serie A", "Eredivisie", "Liga Portugal", "Jupiler Pro League", "Superliga", "Super Lig", "Serie B", "LaLiga2", "Super League", "HNL")
top_competitions

top_competitions_pre_covid_matches_tipsport <- draw_matches_tipsport_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time < "2020-03-01 00:00")
print(paste0("New length of data frame is ", nrow(top_competitions_pre_covid_matches_tipsport)))

top_competitions_pre_covid_matches_fortuna <- draw_matches_fortuna_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time < "2020-03-01 00:00")
print(paste0("New length of data frame is ", nrow(top_competitions_pre_covid_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- top_competitions_pre_covid_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- top_competitions_pre_covid_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="darkorchid", alpha=0.85, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (TOP competitions pre-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_top_competitions_preCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- top_competitions_pre_covid_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- top_competitions_pre_covid_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="darkorchid", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (TOP competitions pre-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_top_competitions_preCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

## Only TOP competitions, only post-COVID seasons

### Filter out low-level competitions and pre-COVID matches too
```{r}
top_countries <- c("England", "Germany", "France", "Czech Republic", "Spain", "Italy", "Netherlands", "Portugal", "Belgium", "Denmark", "Turkey", "Greece", "Switzerland", "Croatia")
top_countries

top_competitions <- c("FORTUNA_LIGA", "Bundesliga", "2. Bundesliga", "Premier League", "Championship", "Ligue 1", "LaLiga", "Serie A", "Eredivisie", "Liga Portugal", "Jupiler Pro League", "Superliga", "Super Lig", "Serie B", "LaLiga2", "Super League", "HNL")
top_competitions

top_competitions_post_covid_matches_tipsport <- draw_matches_tipsport_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time > "2021-09-01 00:00")
print(paste0("New length of data frame is ", nrow(top_competitions_post_covid_matches_tipsport)))

top_competitions_post_covid_matches_fortuna <- draw_matches_fortuna_cut |>
  filter(country %in% top_countries & competition %in% top_competitions
         & date_time > "2021-09-01 00:00")
print(paste0("New length of data frame is ", nrow(top_competitions_post_covid_matches_fortuna)))
```

### Histograms of total goals for each bin of matches
- **Tipsport**
```{r}
summaries <- top_competitions_post_covid_matches_tipsport |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- top_competitions_post_covid_matches_tipsport |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="darkorchid", alpha=0.85, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (TOP competitions post-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_tipsport_top_competitions_postCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```

- **Fortuna**
```{r}
summaries <- top_competitions_post_covid_matches_fortuna |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

chart <- top_competitions_post_covid_matches_fortuna |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="darkorchid", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (TOP competitions post-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")"))

chart

filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\total_goals_in_likely_draw_matches\\", src_basename, "\\histograms_fortuna_top_competitions_postCOVID.png")
ggsave(filename, plot = chart, width = 12, height = 10)
```



- TODO: Maybe based on the variance, find a safe interval for draw odds without margin (for betting)