---
title: "Total_goals_in_likely_draw_matches"
author: "Kutálek Jiří"
date: "2024-02-09"
output:
  html_document:
    theme: default
    smart: no
    toc_depth: 0
    highlight: tango
    df_print: default
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, tidy = FALSE, warning = TRUE, results = 'markup', comment = NA, background = '#FFCC00') # adjust according to your preferences
```

```{r}
library(tidyr)
library(dplyr)
library(agricolae)
library(ggplot2)
library(base)
```

### Read data
```{r}
src_filename = "matches_10_10.csv"

dat_raw <- read.csv2(file = paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\", src_filename), header = TRUE, sep = ",")
```

### Adjust data
- convert to `factors`
- convert `floats` (**Expected Goals** and **Odds**) to `numerics`
- convert other numerical values to `integers`
- convert date & time to `datetime`
- replace conflicting `:` symbol with `_`
- add attribute for **total goals** in match
```{r}
dat <- dat_raw |> mutate(
  across(
    c(team_home, team_away, country, competition, season, referee, neutral_field, finished, no_spectators), factor))
dat <- dat |> mutate(
  across(
    where(is.character) & (starts_with("odd") | starts_with("expected")), as.numeric))
dat <- dat |> mutate(
  across(
    where(is.character) & !c(id, date_time), as.integer))

dat$date_time <- as.POSIXct(dat$date_time, format="%Y-%m-%d %H:%M", tz="UTC")
dat <- dat |> mutate(
  competition = gsub(":", "_", competition)
)

dat <- dat |> mutate(
  total_goals = goals_home + goals_away
)
```

### Add column for **match group** `(country, competition, season)`
```{r}
dat <- dat |> mutate(
    match_group = factor(paste0(country, "_", competition, "_", season)))
```

### Add columns for **margins**
- four columns in total - **start**/**end** ("long time before match start"/"right before amtch start") and **Tipsport**/**Fortuna**
- `NA_real_` if zero or invalid odd values
```{r}
dat <- dat |> mutate(
  margin_tipsport_start = ifelse(
    is.na(odd_tipsport_1_start) | is.na(odd_tipsport_0_start) | is.na(odd_tipsport_2_start) | 
      odd_tipsport_1_start <= 0   | odd_tipsport_0_start <= 0   | odd_tipsport_2_start <= 0,
    NA_real_,
    (1/odd_tipsport_1_start + 1/odd_tipsport_0_start + 1/odd_tipsport_2_start) - 1),
  margin_tipsport_end = ifelse(
    is.na(odd_tipsport_1_end) | is.na(odd_tipsport_0_end) | is.na(odd_tipsport_2_end) | 
      odd_tipsport_1_end <= 0   | odd_tipsport_0_end <= 0   | odd_tipsport_2_end <= 0,
    NA_real_,
    (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end) - 1),
  margin_fortuna_start = ifelse(
    is.na(odd_fortuna_1_start) | is.na(odd_fortuna_0_start) | is.na(odd_fortuna_2_start) | 
      odd_fortuna_1_start <= 0   | odd_fortuna_0_start <= 0   | odd_fortuna_2_start <= 0,
    NA_real_,
    (1/odd_fortuna_1_start + 1/odd_fortuna_0_start + 1/odd_fortuna_2_start) - 1),
  margin_fortuna_end = ifelse(
    is.na(odd_fortuna_1_end) | is.na(odd_fortuna_0_end) | is.na(odd_fortuna_2_end) | 
      odd_fortuna_1_end <= 0   | odd_fortuna_0_end <= 0   | odd_fortuna_2_end <= 0,
    NA_real_,
    (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end) - 1)
)

# Print out average margin (start and end)
print(paste0("Average margin by Tipsport for win/draw/loose is: ",
             mean(dat$margin_tipsport_start, na.rm = TRUE), ", ",
             mean(dat$margin_tipsport_end, na.rm = TRUE)))
print(paste0("Average margin by Fortuna for win/draw/loose is: ",
             mean(dat$margin_fortuna_start, na.rm = TRUE), ", ",
             mean(dat$margin_fortuna_end, na.rm = TRUE)))

last_N_columns = 4
random_rows <- sample(nrow(dat), 10) # randomly take 10 rows
dat[random_rows, c(1, (ncol(dat) - (last_N_columns-1)) : ncol(dat))]
```

### Add columns for win/draw/lose **probabilities without margins** (only **end**, not **start**)
- `prob_without_margin = prob_with_margin - prob_with_margin * (margin / sum_of_odds)`
- (`sum_of_odds = 1 + margin`)
```{r}
dat <- dat |> mutate(
  prb_nomarg_1_tips_end = (1/odd_tipsport_1_end) - (1/odd_tipsport_1_end) * (margin_tipsport_end / (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end)),
  prb_nomarg_0_tips_end = (1/odd_tipsport_0_end) - (1/odd_tipsport_0_end) * (margin_tipsport_end / (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end)),
  prb_nomarg_2_tips_end = (1/odd_tipsport_2_end) - (1/odd_tipsport_2_end) * (margin_tipsport_end / (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end)),
  prb_nomarg_1_fort_end = (1/odd_fortuna_1_end) - (1/odd_fortuna_1_end) * (margin_fortuna_end / (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end)),
  prb_nomarg_0_fort_end = (1/odd_fortuna_0_end) - (1/odd_fortuna_0_end) * (margin_fortuna_end / (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end)),
  prb_nomarg_2_fort_end = (1/odd_fortuna_2_end) - (1/odd_fortuna_2_end) * (margin_fortuna_end / (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end))
)

# Print out average probabilities without margins (end, not start)
print(paste0("Average probabilities without margins by Tipsport for win/draw/loose are: ",
             mean(dat$prb_nomarg_1_tips_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_0_tips_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_2_tips_end, na.rm = TRUE)))
print(paste0("Average probabilities without margins by Fortuna for win/draw/loose are: ",
             mean(dat$prb_nomarg_1_fort_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_0_fort_end, na.rm = TRUE), ", ",
             mean(dat$prb_nomarg_2_fort_end, na.rm = TRUE)))

last_N_columns = 6
random_rows <- sample(nrow(dat), 10) # randomly take 10 rows
dat[random_rows, c(1, (ncol(dat) - (last_N_columns-1)) : ncol(dat))]
```

### Filter and check the calculated probabilities
- filter out matches with `NA` values for probabilities
- check that the probabilities for each match summing to `1`
```{r}
dat_filtered <- dat |> filter(!is.na(prb_nomarg_1_tips_end) & !is.na(prb_nomarg_0_tips_end) & !is.na(prb_nomarg_2_tips_end) & !is.na(prb_nomarg_1_fort_end) & !is.na(prb_nomarg_0_fort_end) & !is.na(prb_nomarg_2_fort_end))
print(paste0("New length of data frame is ", nrow(dat_filtered)))

# Check the probabilities
prob_columns <- dat_filtered[c("prb_nomarg_1_tips_end", "prb_nomarg_0_tips_end", "prb_nomarg_2_tips_end", "prb_nomarg_1_fort_end", "prb_nomarg_0_fort_end", "prb_nomarg_2_fort_end")]

tolerance <- 1e-5
rows_outside_range <- apply(prob_columns, 1, function(x) {
  prob_sum <- sum(x, na.rm = TRUE)
  abs(prob_sum - 1) > tolerance
})

print(paste0("Number of wrongly calculated matches probabilities equals to ", sum(rows_outside_range, na.rm = TRUE)))
rows_indices_outside_range <- which(rows_outside_range)
```

### Add column for **absolute difference** in **winning probability** (without margin) for both teams
```{r}
dat_filtered <- dat_filtered |> mutate(
  diff_prb_nomarg_tips_end = abs(prb_nomarg_1_tips_end - prb_nomarg_2_tips_end),
  diff_prb_nomarg_fort_end = abs(prb_nomarg_1_fort_end - prb_nomarg_2_fort_end)
)

last_N_columns = 2
random_rows <- sample(nrow(dat_filtered), 10) # randomly take 10 rows
dat_filtered[random_rows, c(1, (ncol(dat_filtered) - (last_N_columns-1)) : ncol(dat_filtered))]
```

### **Filter out** matches during the COVID-19 period and those without spectators
```{r}
no_spectators_matches <- dat_filtered |> filter(no_spectators == "True")
print(paste0("New length of data frame is ", nrow(dat_filtered)))

dat_filtered <- dat_filtered |> filter(no_spectators == "False" & (date_time < "2020-03-01 00:00" | date_time > "2021-09-01 00:00"))
print(paste0("New length of data frame is ", nrow(dat_filtered)))
```

### **Selection** of possible draw matches
```{r}
draw_matches_tipsport <- dat_filtered |> filter(prb_nomarg_0_tips_end >= 0.28 &
                                     prb_nomarg_0_tips_end <= 0.35 &
                                     diff_prb_nomarg_tips_end <= 0.15)
draw_matches_fortuna <- dat_filtered |> filter(prb_nomarg_0_fort_end >= 0.28 &
                                     prb_nomarg_0_fort_end <= 0.35 &
                                     diff_prb_nomarg_fort_end <= 0.15)

print(paste0("New length of Tipsport data frame is ", nrow(draw_matches_tipsport)))
print(paste0("New length of Fortuna data frame is ", nrow(draw_matches_fortuna)))

last_N_columns = 8
random_rows <- sample(nrow(draw_matches_tipsport), 10) # randomly take 10 rows
draw_matches_tipsport[random_rows, c(1, (ncol(draw_matches_tipsport) - (last_N_columns-1)) : ncol(draw_matches_tipsport))]

last_N_columns = 8
random_rows <- sample(nrow(draw_matches_fortuna), 10) # randomly take 10 rows
draw_matches_fortuna[random_rows, c(1, (ncol(draw_matches_fortuna) - (last_N_columns-1)) : ncol(draw_matches_fortuna))]
```

### Distribute all the selected matches into bins
- as follows: `[2.7-2.8, 2.8-2.9, ..., 3.4-3.5]`
```{r}
draw_matches_tipsport_cut <- draw_matches_tipsport |> mutate(
  bin = cut(prb_nomarg_0_tips_end, 
            breaks = c(0.28, 0.29, 0.30, 0.31, 0.32, 0.33, 0.34, 0.35),
            labels = c('28-29%', '29-30%', '30-31%', '31-32%', '32-33%', '33-34%', '34-35%'),
            include.lowest = TRUE, right = FALSE)
)
table(draw_matches_tipsport_cut$bin)

draw_matches_fortuna_cut <- draw_matches_fortuna |> mutate(
  bin = cut(prb_nomarg_0_fort_end, 
            breaks = c(0.28, 0.29, 0.30, 0.31, 0.32, 0.33, 0.34, 0.35),
            labels = c('28-29%', '29-30%', '30-31%', '31-32%', '32-33%', '33-34%', '34-35%'),
            include.lowest = TRUE, right = FALSE)
)
table(draw_matches_fortuna_cut$bin)
```

# Histograms of total goals for each bin of matches
```{r}
summaries <- draw_matches_tipsport_cut |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

draw_matches_tipsport_cut |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (all non-COVID matches with spectators; Tipsport)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename)) +
  theme(plot.title = element_text(size = 9))


summaries <- draw_matches_fortuna_cut |>
  group_by(bin) |>
  summarize(mean_goals = mean(total_goals, na.rm=T),
            variance = var(total_goals, na.rm=T),
            max_count = max(table(total_goals)),  # the highest bar for positioning
            total_matches = n())  # total number of matches in each bin

draw_matches_fortuna_cut |> ggplot(aes(x=total_goals)) + 
  geom_histogram(binwidth=1, fill="blue", alpha=0.7, color="white") +
  facet_wrap(~bin, scales="free_y") + 
  geom_vline(data=summaries, aes(xintercept=mean_goals),
             color="red", linetype="dashed", size=1) + 
  geom_text(data=summaries, 
            aes(x=mean_goals + 1, y=max_count - 1, 
                label=sprintf("Mean: %.2f\nVar: %.2f\nMatches: %d", mean_goals, variance, total_matches)), 
            hjust=0, vjust=1) +
  labs(title="Histogram of total goals in likely draw resulting matches (all non-COVID matches with spectators; Fortuna)",
       x="Total Goals", y=paste0("Number of Matches (based on ", src_filename, ")")) +
  theme(plot.title = element_text(size = 9))
```