---
title: "margins_odds_per_comp_season"
author: "Kutálek Jiří"
date: "2024-02-09"
output: 
  html_document:
    theme: default
    smart: no
    toc_depth: 0
    highlight: tango
    df_print: default
    code_folding: show
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, tidy = FALSE, warning = TRUE, results = 'markup', comment = NA, background = '#FFCC00') # adjust according to your preferences
```

```{r}
library(tidyr)
library(dplyr)
library(agricolae)
library(ggplot2)
library(base)
```

### Read data
```{r}
src_filename = "matches_10_10.csv"

dat_raw <- read.csv2(file = paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\", src_filename), header = TRUE, sep = ",")
```

### Adjust data
- convert to `factors`
- convert `floats` (**Expected Goals** and **Odds**) to `numerics`
- convert other numerical values to `integers`
```{r}
dat <- dat_raw |> mutate(
  across(
    c(team_home, team_away, country, competition, season, referee, neutral_field, finished, no_spectators), factor))
dat <- dat |> mutate(
  across(
    where(is.character) & (starts_with("odd") | starts_with("expected")), as.numeric))
dat <- dat |> mutate(
  across(
    where(is.character) & !c(id, date_time), as.integer))
```

### Add column for **match group** `(country, competition, season)`
- one violin plot per match group
```{r}
dat <- dat |> mutate(
    match_group = factor(paste0(country, "_", competition, "_", season)))
```

### Replace the `:` symbol for **FORTUNA:LIGA** match group by `_`
```{r}
dat$match_group <- gsub(":", "_", dat$match_group)
```

### Add columns for **margins**
- four columns in total - **start**/**end** ("long time before match start"/"right before amtch start") and **Tipsport**/**Fortuna**
- `NA_real_` if zero or invalid odd values
```{r}
dat <- dat |> mutate(
  margin_tipsport_start = ifelse(
    is.na(odd_tipsport_1_start) | is.na(odd_tipsport_0_start) | is.na(odd_tipsport_2_start) | 
      odd_tipsport_1_start <= 0   | odd_tipsport_0_start <= 0   | odd_tipsport_2_start <= 0,
    NA_real_,
    (1/odd_tipsport_1_start + 1/odd_tipsport_0_start + 1/odd_tipsport_2_start) - 1),
  margin_tipsport_end = ifelse(
    is.na(odd_tipsport_1_end) | is.na(odd_tipsport_0_end) | is.na(odd_tipsport_2_end) | 
      odd_tipsport_1_end <= 0   | odd_tipsport_0_end <= 0   | odd_tipsport_2_end <= 0,
    NA_real_,
    (1/odd_tipsport_1_end + 1/odd_tipsport_0_end + 1/odd_tipsport_2_end) - 1),
  margin_fortuna_start = ifelse(
    is.na(odd_fortuna_1_start) | is.na(odd_fortuna_0_start) | is.na(odd_fortuna_2_start) | 
      odd_fortuna_1_start <= 0   | odd_fortuna_0_start <= 0   | odd_fortuna_2_start <= 0,
    NA_real_,
    (1/odd_fortuna_1_start + 1/odd_fortuna_0_start + 1/odd_fortuna_2_start) - 1),
  margin_fortuna_end = ifelse(
    is.na(odd_fortuna_1_end) | is.na(odd_fortuna_0_end) | is.na(odd_fortuna_2_end) | 
      odd_fortuna_1_end <= 0   | odd_fortuna_0_end <= 0   | odd_fortuna_2_end <= 0,
    NA_real_,
    (1/odd_fortuna_1_end + 1/odd_fortuna_0_end + 1/odd_fortuna_2_end) - 1)
)
```

##### Descriptive statistics for margins in each match group
- *just example, not needed*
```{r}
dat |> group_by(match_group) |>
  summarise(
    across(
      margin_tipsport_start, list(
        mean = ~mean(., na.rm = TRUE),
        sd = ~sd(., na.rm = TRUE),
        var = ~var(., na.rm = TRUE),
        median = ~median(., na.rm = TRUE),
        IQR = ~IQR(., na.rm = TRUE)
    ), .names = "{.fn}")
  ) |> as.data.frame()
# ...dtto for margin_tipsport_end, margin_fortuna_start and margin_fortuna_end...
```

##### Sample visualization of margins for a match group
```{r}
dat |> filter(match_group == "Argentina_Liga Profesional_2018-2019") |>
  pivot_longer(
    cols = c(margin_tipsport_start, margin_tipsport_end, margin_fortuna_start, margin_fortuna_end),
    names_to = "company_startend",
    values_to = "margin") |> 
  mutate(company_startend = factor(company_startend, levels = c("margin_tipsport_start", "margin_tipsport_end", "margin_fortuna_start", "margin_fortuna_end"))) |>
  ggplot(aes(x = company_startend, y = margin, fill = company_startend)) + 
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.125, ymax = 0.100), fill = "#ffbfbf", alpha = 0.2) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.100, ymax = 0.075), fill = "#ffdebf", alpha = 0.2) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.075, ymax = 0.050), fill = "#fffabf", alpha = 0.2) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.050, ymax = 0.025), fill = "#e5ffbf", alpha = 0.2) +
  geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.025, ymax = 0.000), fill = "#c5efbf", alpha = 0.2) +
  geom_violin() +
  geom_jitter(width = 0.2, alpha = 0.4, size = 0.2) +
  labs(title = "Margins in matches of Argentina_Liga Profesional_2018-2019",
       x = "Betting company + Start/End of odds publishing period",
       y = "Margin") +
  coord_cartesian(ylim = c(0.0, 0.12)) +
  scale_fill_manual(values = c("lightgoldenrod1", "yellow1", "lightskyblue1", "dodgerblue"))
```

### Complete and save visualization of margins for all match groups
```{r}
all_match_groups <- factor(unique(dat$match_group))

for(group in all_match_groups) {
  data <- dat |> filter(match_group == group) |>
    pivot_longer(
      cols = c(margin_tipsport_start, margin_tipsport_end, margin_fortuna_start, margin_fortuna_end),
      names_to = "company_startend",
      values_to = "margin") |>
    mutate(company_startend = factor(company_startend, levels = c("margin_tipsport_start", "margin_tipsport_end", "margin_fortuna_start", "margin_fortuna_end")))  # correct order of these 4 groups
  
  chart <- data |> ggplot(aes(x = company_startend, y = margin, fill = company_startend)) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.125, ymax = 0.100), fill = "#ffbfbf", alpha = 0.2) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.100, ymax = 0.075), fill = "#ffdebf", alpha = 0.2) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.075, ymax = 0.050), fill = "#fffabf", alpha = 0.2) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.050, ymax = 0.025), fill = "#e5ffbf", alpha = 0.2) +
    geom_rect(aes(xmin = -Inf, xmax = Inf, ymin = 0.025, ymax = 0.000), fill = "#c5efbf", alpha = 0.2) +
    geom_violin() +
    geom_jitter(width = 0.2, alpha = 0.4, size = 0.2) +
    labs(title = paste("Margins in matches of ", group),
         x = "Betting company + Start/End of odds publishing period",
         y = paste0("Margin (based on )", src_filename)) +
    coord_cartesian(ylim = c(0.0, 0.12)) +
    scale_fill_manual(values = c("lightgoldenrod1", "yellow1", "lightskyblue1", "dodgerblue"))

  filename <- paste0("C:\\Users\\kutalekj\\PycharmProjects\\MyFlashscoreScraper\\src\\margins_violinplots\\Margins_", gsub(" ", "_", group), ".png")

  ggsave(filename, plot = chart, width = 12, height = 10)
}
```